---
title: "Lab 1, Winter 2021 (1/8/21)"
author: "Dylan Glave"
date: "1/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```




## Lab #1, Winter 2020: Wrangling and Maps (of San Francisco, no less!)


### First, I'll just read in the map of SF streets
```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic Wrangling Reminders
I'm refreshing some skills for wrangling and summary statistics. this is from 'dplyr' package

Let's find the Top 5 highest observations of trees, by legal status. Then I'll wrangle. Then I'll make a graph. 
```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  ### Select would work to select or remove columns, we'll use relocate! Relocate can move around any columns or variables. We put jsut 1 variable name, which moved that variable to the "front" (the left)
relocate(tree_count) %>% 
  slice_max(tree_count, n = 5) 
```

Make a graph of the top 5 observations, by legal status. This is using the wrangled version of sf_trees, called top_5_status

```{r}
ggplot(data = top_5_status, aes (x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col() + 
  labs( x = "Legal Status", y = "Tree Count", title = "SF Trees") +
  coord_flip() +
  theme_minimal()
```

### A few more Wrangling examples. I only want to keep observations (rows) of blackwood acacia trees. 

```{r}
blackwood_acacia <-  sf_trees %>% 
filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)
```

Making a faux-map, based on Latitude and Longitude. BUT R does not know that Lat/Long are spatial coordinates. 

```{r}
ggplot(data = blackwood_acacia, aes( x = longitude, y = latitude)) +
  geom_point( size = 0.5)
```


### Using tidyr::separate() and unite() functions. 

Useful for combining or separating columns. 

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

Example: tidyr::unite to unite tree and legal status (weird)

```{r}
sf_tree_united <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")
```




